One Year HPC Certification Forum in Retrospective

### 2019-04-17 paper.tex
edited the Overview of the actities section

The ever-changing nature of HPC has always compelled the HPC community to focus a lot of effort into training of new and existing practitioners. Historically, these efforts were tailored
 around a typical group of users possessing, due to their background, a certain set of programming skills. However, as HPC has become more diverse in terms of hardware, software and
 the user background, the traditional training approaches became insufficient in addressing training needs of our community. This increasingly complicated HPC landscape makes
 development and delivery of new training materials challenging. How should we develop training for users, often coming from non-traditionally HPC disciplines, and only interested in
 learning a particular set of skills? How can we satisfy their training needs if we don't really understand what these are? It's clear that HPC centres struggle to identify and overcome the
 gaps in users' knowledge, while users struggle to identify skills required to perform their tasks.

@@

Most institutions operating HPC systems offer regular training events focusing on general aspects of their supercomputer's hardware architecture, software stack, application development environment  and various tuning and debugging tools.
It is understandable why the materials they are geared towards the special demands of the institutions they support and its specific HPC environment.
The problem is that teaching content typically covers only a small part of basic HPC skills necessary to use other HPC systems.
This approach of teaching only specific implementations, tools or workflows, instead of concepts behind them makes a move to a different system or tool unnecessary complicated.
Narrowing the scope of training events make sense from the HPC provider perspective, but is not very conducive to development of more comprehensive learning environment that provides assessment and certifies the newly acquired skills. So although, certificates are used widely in IT industry to verify certain knowledge, until now there was no similar approach for HPC training.
It is, however, clear that a mapping of competences and certification scheme could address some of the challenges stemming from the constantly growing training needs of our community,

@@

During the first year of the existence of the HPCCF, the discussion within the forum and with partners has lead to the establishment of various processes and prototypes for technical tools that are described in the following briefly.
Further details of the overview are provided in subsequent sections.

\paragraph{Management}
At ISC-HPC 2018, the first steering board had been elected.
We established communication channels via webpage, email and Slack for regular discussion.
Slack is also used to conduct our monthly open meetings between members and the steering board.
At beginning we used video conferencing tools, but found that the chat provides a more inclusive discussion and the asynchronous nature is more productive.
We also established relations with various stakeholders involved in HPC education growing the participation in the forum.
Meeting notes with action items are extracted from Slack and documented using Google Doc\kh{Vorschlag: kurze Info zum Tool bzw. Link (wie bei Slack) ergänzen}.

Note that until now contributions of members are voluntarily and follow the schedule as agreed between the contributor and the HPCCF board.
We manage the prospect contributions in a Trello board\kh{Vorschlag: kurze Info zum Tool bzw. Link (wie bei Slack und Google Doc) ergänzen (Trello kennt ja nicht unbedingt jeder Leser); der Begriff "collaborative" sollte irgendwo untergebracht werden}; each member can manage its own card.

\paragraph{Technical}\kh{"Management" (oberhalb) finde ich auf der Ebene alleinstehend gut, "Technical" etwas zu kurz, Vorschlag: "Technical Tasks" (o.ä.)}
According to our goals, we address 1) the definition of competencies; 2) the creation of certificates; 3) the examination of practitioners; and 4) an ecosystem of tools supporting them.

For all these sub-goals we made substantial progress.
While the definition and organization of competencies was the main focus, we prototyped various tools and processes that embed the competencies into the wider education ecosystem.

Firstly, we finalized the templates for the description of a competence considering pedagogic literature from higher education.
We are still working on mapping out all relevant competencies, but included also new topics that will be covered in the first version of the certification program.

The definition of the competencies covered by the program is version controlled with Git.
The competencies are available in XML and Markdown format.
A Wiki allows to edit the competencies and simplifies access for non-technical users.
A navigable JavaScript representation of the skill tree is available that can be customized for the HPCCF but also for end-users.

Prototypical processes and tools have been developed and deployed for curation of the examination questions and for conducting an online multiple-choice examination.

>>>>

The ever-changing nature of HPC has always compelled the HPC community to focus a lot of effort into training of new and existing practitioners. Historically, these efforts were tailored
 around a typical group of users possessing, due to their background, a certain set of programming skills. However, as HPC has become more diverse in terms of hardware, software and 
 the user background, the traditional training approaches became insufficient in addressing training needs of our community. This increasingly complicated HPC landscape makes 
 developement and delivery of new training materials challenging. How should we develop training for users, often coming from non-traditionally HPC disciplines, and only interested in 
 learning a particular set of skills? How can we satisfy their training needs if we don't really understand what these are? It's clear that HPC centres struggle to identify and overcome the 
 gaps in users' knowledge, while users struggle to identify skills required to perform their tasks.

@@

Most institutions operating HPC systems offer regular training events focusing on general aspects of their supercomputer's hardware architecture, software stack, application development environment  and various tuning and debugging tools.   
It is understandable why the materials they are geared towards the special demands of the institutions they support and its specific HPC environment. 
The problem is that teaching content typically covers only a small part of basic HPC skills necessary to use other HPC systems. 
This approach of teaching only specific implementations, tools or workflows, instead of concepts behind them makes a move to a different system or tool unnecessary complicated. 
Narrowing the scope of training events make sense from the HPC provider perspective, but is not very conducive to development of more comprehensive learning environment that provides assessment and certifies the newly acquired skills. So although, certificates are used widely in IT industry to verify certain knowledge, until now there was no similar approach for HPC training. 
It is, however, clear that a certification scheme could address some of the challenges stemming from the constantly growing training needs of our community,

@@

The first year of the HPCCF forum was rich in discussions, both internally and with external partners, resulting in a number of processes and prototype tools being put in place. These are described briefly below, and in more details in the subsequent sections.

\paragraph{Management}
The first steering board has been elected at ISC-HPC 2018.
We established communication channels via webpage, email and Slack for regular discussion.
Slack is also used to conduct our monthly open meetings between members and the steering board.
At beginning video conferencing tools have been used, but the geographical diversity of the attendees influenced the decision to move towards chat-based tools. 
The asynchronous nature and automatic message retention proved to be more productive, and provided a more open and inclusive platform for discussion.    
Meeting notes with action items are extracted from Slack and documented using Google Doc (an online word processor allowing a real-time and multiple-user collaboration).

The ongoing conversation with various stakeholders involved in HPC education and training, keeps producing more engagement in the forum's activities.   
The member's contributions are voluntarily and follow the schedule agreed between the contributor and the HPCCF board.
The prospective contributions are managed in Trello (a web-based list-making collaborative application), in which each member can manage their own card.

\paragraph{Technical Tasks}
The tasks we set out to accomplish are:  1) the definition of competencies; 2) the examination of practitioners;  3) the creation of certificates; and 4) the reinforcement of an ecosystem of tools supporting them.

For all these sub-goals we made substantial progress.
While the definition and organisation of competencies was the main focus, we prototyped various tools and processes that embed the competencies into the wider education ecosystem.

Firstly, considering the pedagogic literature from higher education we finalised the templates for the skills description. 
The effort to map out all of the relevant competencies is still on-going, but the feedback received during a number of events organised by the HPCCF and other members of the HPC education community resulted in the addition of new topics to the first version of the certification program.

The definitions of the competencies are version controlled with Git, and are publicly available in XML and Markdown format.
A Wiki\footnote{https://www.hpc-certification.org/wiki/} is being used to edit the skills' descriptions, and a navigable and widely customisable JavaScript\footnote{https://www.hpc-certification.org/skills/map/} makes the skill tree easily accessible to the forum members and the end-users alike.

Premilinary processes and tools have been developed and deployed for curation of the examination questions and for conducting an online multiple-choice examination.
